---
title: "Overall Effect of N"
author: "Ursula Ruiz-Vera, Deepak Jaiswal, David LeBauer"
date: "Fri Nov 07 16:04:51 2014"
output:
  html_document: default
  pdf_document:
    keep_tex: yes
    number_sections: yes
  word_document: default
---
```{r }
#setwd("C:\\Users\\ruizver1\\Documents\\myprojectEBI\\sugarbag")
library(sugarbag)
library(data.table)
library(ggplot2)
setwd("./")
knitr::opts_chunk$set(warning=FALSE, cache=TRUE)
```
## Prepare data
```{r}
# Step 1:- Get all the experiment IDs  for whic Harvest, Experiment Design and Planting information is available

expID <-GetExpID_for_N_Combined()
# Step 2 :- Define variable name that we are interested in

varname <- c("DWMST","DWTOTA","LAI",  "NWTOTA", "SWMST","SWTOTA", "Latitude.x", "Longitude.x")

## Create dataframe
id <- expID[1]
Output <- Update_expID(id, varname)
N = length(expID)

# Get N versus Yield data from all the experiments andRBIND them to Output
for (i in 2:N){
  id <- expID[i]
  tmp_output <- Update_expID(id, varname)
  Output <- rbind(Output,tmp_output)
}
```
## Calculate the plot means and then the maximum values of the variables
```{r}

OUT <- data.table(Output)

OUT2 <- OUT[,list(DWMST=mean(DWMST, na.rm = TRUE),DWTOTA=mean(DWTOTA, na.rm = TRUE),LAI=mean(LAI, na.rm = TRUE), 
                  NWTOTA=mean(NWTOTA, na.rm = TRUE),SWMST=mean(SWMST, na.rm = TRUE),SWTOTA=mean(SWTOTA, na.rm = TRUE)), 
            by = c('ExpID', 'Treatment', 'Date', 'Ncombined', 'Cultivar', 'Irrigated')]


ndata <- unique(OUT2[,list(DWMST=max(DWMST, na.rm = TRUE),DWTOTA=max(DWTOTA, na.rm = TRUE),LAI=max(LAI, na.rm = TRUE), 
                  NWTOTA=max(NWTOTA, na.rm = TRUE),SWMST=max(SWMST, na.rm = TRUE),SWTOTA=max(SWTOTA, na.rm = TRUE)),
            by = c('ExpID','Treatment','Ncombined','Cultivar','Irrigated')])

d <- ndata[!is.na(Ncombined), 
           list(ExpID, Treatment, Ncombined = as.numeric(Ncombined), Cultivar, Irrigated, DWTOTA, SWTOTA, NWTOTA, DWMST, SWMST, 
          yield = DWTOTA, sugar = SWTOTA, cane_yield = DWMST, cane_sugar = SWMST/DWMST, nitrogen_concentration = NWTOTA/DWTOTA)]

## Clean -Inf http://stackoverflow.com/a/12188551/199217
invisible(lapply(names(d),function(.name) set(d, which(is.infinite(d[[.name]])), j = .name,value =NA)))
```
# Meta-analysis of N effects on yield, Sugar, and LAI

## Exploratory Data Analysis

Our question is:
* Is sugarcane yield and sugar content limited by N and/or water availablility?
Our Model: 
* y = Nitrogen * Irrigation + epsilon
where y is yield or sugar content (of total aboveground biomass, of cane)

### Summarize
```{r}
library(pander)
#Records where Cane Sugar Content are too low, it is because in some data DWMST is NA and it is giving cane_sugar=0
pander(d[cane_sugar < 0.1 & cane_sugar > 0])
d[cane_sugar < 0.1,`:=` (cane_sugar = NA)]

pander(data.frame(n = nrow(d)))
pander(summary(d[, list(Ncombined, yield, sugar, cane_yield, cane_sugar, nitrogen_concentration)], digits = 3))
pander(summary(lm(yield ~ Ncombined*Irrigated, data = d), digits = 3))
pander(summary(lm(cane_yield ~ Ncombined*Irrigated, data = d), digits = 3))
pander(summary(lm(cane_sugar ~ Ncombined*Irrigated, data = d), digits = 3))
```
### Check distribution, outliers, odd variables 
```{r}
ggplot(data = d) + geom_histogram(aes(yield))

ggplot(data = d) + geom_histogram(aes(sugar))

ggplot(data = d) + geom_histogram(aes(cane_yield))

ggplot(data = d) + geom_histogram(aes(cane_sugar))
```
## Look which ExpID had irrigation and non-irrigation
```{r}
data(variables)
if(exists("variables"))pander(variables[variables$VariableName %in% colnames(ndata),])

ggplot(data = d) + geom_point(aes(Ncombined, yield, color = as.logical(Irrigated))) + 
  facet_wrap(~ExpID) + ylim(range(0, max(pretty(ndata$DWTOTA))))

ggplot(data = d) + geom_point(aes(Ncombined, sugar, color = as.logical(Irrigated))) + 
  facet_wrap(~ExpID) + ylim(range(0, 4000))

ggplot(data = d) + geom_point(aes(Ncombined, cane_yield, color = as.logical(Irrigated))) + 
  facet_wrap(~ExpID) + ylim(range(0, 8000))

ggplot(data = d) + geom_point(aes(Ncombined, cane_sugar, color = as.logical(Irrigated))) + 
  facet_wrap(~ExpID) + ylim(range(0, 0.6))

ggplot(data = d) + geom_point(aes(Ncombined, nitrogen_concentration, color = as.logical(Irrigated))) + 
  facet_wrap(~ExpID) + ylim(range(0, 0.008))
```

```{r}
##graphic with linear regression (by Ursula)

ggplot(data = d, aes(Ncombined, yield)) + 
  geom_point(aes(Ncombined, yield, color = Cultivar), alpha = 0.4) + 
  stat_smooth(method=lm, level=0.9) +
  facet_wrap(~Irrigated) + ylim(0,10000) + 
  xlab("amount of N (Kg/ha)") + ylab("Dry Weight of Above-ground Biomass (g/m2)") 

ggplot(data = d, aes(Ncombined, sugar)) + 
  geom_point(aes(Ncombined, sugar, color = Cultivar), alpha = 0.4) + 
  facet_wrap(~Irrigated) + 
  stat_smooth(method=lm, level=0.9) +ylim(0,4000) + 
   xlab("amount of N (Kg/ha)") + ylab("Sucrose in Above-ground Biomass (g/m2)")

ggplot(data = d, aes(Ncombined, cane_yield)) + 
  geom_point(aes(Ncombined, cane_yield, color = Cultivar), alpha = 0.4) + 
  stat_smooth(method=lm, level=0.9) +
  facet_wrap(~Irrigated)+  ylim(0,8000) +
  xlab("amount of N (Kg/ha)") + ylab("Dry weight of the Cane Biomass (g/m2)")

ggplot(data = d, aes(Ncombined, cane_sugar)) + 
  geom_point(aes(Ncombined, cane_sugar, color = Cultivar), alpha = 0.4) + 
  facet_wrap(~Irrigated) +  
  stat_smooth(method=lm, level=0.9) + ylim(0,0.6) + 
  xlab("amount of N (Kg/ha)") + 
  ylab("Sucrose/Dry Weight, of the Cane Biomass")

ggplot(data = d, aes(Ncombined,nitrogen_concentration) ) + 
  geom_point(aes(Ncombined, nitrogen_concentration, color = Cultivar), alpha = 0.4) + 
  facet_wrap(~Irrigated) + 
  stat_smooth(method=lm, level=0.9) + ylim(0,0.008) + 
  xlab("amount of N (Kg/ha)") + ylab("Nitrogen/Dry Weight, of the Above-ground Biomass")
####
```
### Check basic models (testing normality with shapiro test and transforminng some variables, linear regression analysis (modified by Ursula))
```{r}
#test
par(mfrow=c(2,2))

library(nlme)
install.packages("lme4")

hist(d$Ncombined)
hist(sqrt(d$Ncombined))

shapiro.test(d$yield)
shapiro.test(log10(d$yield))
shapiro.test(log(d$yield))
shapiro.test(sqrt(d$yield))

d[,plot(Ncombined, log10(yield))]
d[,plot(sqrt(Ncombined), log10(yield))]

plot(lm(log10(yield) ~ Ncombined * as.factor(Irrigated), data = d), main = "Yield")
summary(lm(log10(yield) ~ Ncombined * as.factor(Irrigated), data = d), main="Yield")
#plot(lm(log10(yield) ~ sqrt(Ncombined) * as.factor(Irrigated), data = d), main = "Yield")
#summary(lm(log10(yield) ~ sqrt(Ncombined) * as.factor(Irrigated), data = d), main = "Yield")
```

```{r}
#with other variables
shapiro.test(d$sugar)
shapiro.test(log10(d$sugar))
shapiro.test(log(d$sugar))
shapiro.test(sqrt(d$sugar))
hist(d$sugar)
plot(lm(sugar ~ Ncombined * as.factor(Irrigated), data = d), main = "Sugar")
summary(lm(sugar ~ Ncombined * as.factor(Irrigated), data = d), main = "Sugar")
#plot(lm(sugar ~ sqrt(Ncombined) * as.factor(Irrigated), data = d), main = "Sugar")
#summary(lm(sugar ~ sqrt(Ncombined) * as.factor(Irrigated), data = d), main = "Sugar")

shapiro.test(d$cane_yield)
shapiro.test(log10(d$cane_yield))
shapiro.test(log(d$cane_yield))
shapiro.test(sqrt(d$cane_yield))
hist(d$cane_yield)
hist(log10(d$cane_yield))
plot(lm(log10(cane_yield) ~ Ncombined * as.factor(Irrigated), data = d), main = "Cane Yield")
summary(lm(log10(cane_yield) ~ Ncombined * as.factor(Irrigated), data = d), main = "Cane Yield")
#plot(lm(log10(cane_yield) ~ sqrt(Ncombined) * as.factor(Irrigated), data = d), main = "Cane Yield")
#summary(lm(log10(cane_yield) ~ sqrt(Ncombined) * as.factor(Irrigated), data = d), main = "Cane Yield")

#this one (cane_sugar) doesn't fit in any transformation criteria
shapiro.test(d$cane_sugar)
shapiro.test(log10(d$cane_sugar))
shapiro.test(log(d$cane_sugar))
shapiro.test(sqrt(d$cane_sugar))
hist(d$cane_sugar)
plot(lm(cane_sugar ~ Ncombined * as.factor(Irrigated), data = d), main = "Cane % Sugar")
summary(lm(cane_sugar ~ Ncombined * as.factor(Irrigated), data = d), main = "Cane % Sugar")
#plot(lm(cane_sugar ~ sqrt(Ncombined) * as.factor(Irrigated), data = d), main = "Cane % Sugar")
#summary(lm(cane_sugar ~ sqrt(Ncombined) * as.factor(Irrigated), data = d), main = "Cane % Sugar")

shapiro.test(d$nitrogen_concentration)
shapiro.test(log10(d$nitrogen_concentration))
shapiro.test(log(d$nitrogen_concentration))
shapiro.test(sqrt(d$nitrogen_concentration))
hist(d$nitrogen_concentration)
plot(lm(nitrogen_concentration ~ Ncombined * as.factor(Irrigated), data = d), main = "% Nitrogen")
summary(lm(nitrogen_concentration ~ Ncombined * as.factor(Irrigated), data = d), main = "% Nitrogen")
#plot(lm(nitrogen_concentration ~ sqrt(Ncombined) * as.factor(Irrigated), data = d), main = "% Nitrogen")
#summary(lm(nitrogen_concentration ~ sqrt(Ncombined) * as.factor(Irrigated), data = d), main = "% Nitrogen")
```
## Regression Models
#These are throwing an error ... still need to debug.
### Total Aboveground Biomass
```{r}
summary(nlme::lme(fixed = log(yield) ~ Ncombined * as.factor(Irrigated), random = ~1|ExpID, data = d, na.action = na.omit))
#summary(nlme::lme(fixed = log(yield) ~ sqrt(Ncombined) * as.factor(Irrigated), random = ~1|ExpID, data = d, na.action = na.omit))
```
## Total Sugar Content
```{r}
summary(nlme::lme(fixed = sugar ~ Ncombined * as.factor(Irrigated), random = ~1|ExpID, data = d, na.action = na.omit))
#summary(nlme::lme(fixed = sugar ~ sqrt(Ncombined) * as.factor(Irrigated), random = ~1|ExpID, data = d, na.action = na.omit))
```
### Cane_Yield
```{r}
summary(nlme::lme(fixed = log(cane_yield) ~ Ncombined * as.factor(Irrigated), random = ~1|ExpID, data = d, na.action = na.omit))
summary(nlme::lme(fixed = log(cane_yield) ~ sqrt(Ncombined) * as.factor(Irrigated), random = ~1|ExpID, data = d, na.action = na.omit))
```
### Sugar Content of Cane
```{r}
summary(nlme::lme(fixed = cane_sugar ~ Ncombined * as.factor(Irrigated), random = ~1|ExpID, data = d, na.action = na.omit))
#summary(nlme::lme(fixed = cane_sugar ~ sqrt(Ncombined) * as.factor(Irrigated), random = ~1|ExpID, data = d, na.action = na.omit))
```
## Nitrogen Concentration
```{r}
summary(nlme::lme(fixed = nitrogen_concentration ~ Ncombined * as.factor(Irrigated), random = ~1|ExpID, data = d, na.action = na.omit))
#summary(nlme::lme(fixed = nitrogen_concentration ~ sqrt(Ncombined) * as.factor(Irrigated), random = ~1|ExpID, data = d, na.action = na.omit))
```

